---
title: "LastName-FirstName-Project3.Rmd"
# output: html_document
output: pdf_document
---


```{r}
# install.packages("pacman") 
cat("\014") # clears console 
rm(list = ls()) # clears global environment 
try(dev.off(dev.list()["RStudioGD"]), silent = TRUE) # clears plots 
try(p_unload(p_loaded(), character.only = TRUE), silent = TRUE) # clears packages 
options(scipen = 100) # disables scientific notation for entire R session

# You should do this line only once in the entire course. 
# install.packages("pacman") 

library(pacman)
p_load(testthat) 
p_load(tidyverse)
p_load(lubridate)
p_load(ggplot2)
p_load(janitor)

```
# Develop 3–4 data questions to guide the exploratory data analysis
```{r}
# 1. What is the distribution of app ratings across different  categories?
# 2. How does the number of installs correlate with the app rating?
# 3. What are the most common content ratings for apps in different categories?
# 4. How do free apps compare with paid apps in terms of average rating and number of installs?

```

# Part I – Exploring

```{r}
# Data Reading and Initial Exploration

# Reads the dataset
google_play_store = read.csv("google_play_store.csv")

# Prints
google_play_store
```


```{r}
# 1. Review any written description of your dataset. This is often referenced as the data dictionary.

#The dataset is about apps available on the Google Play Store. Each row represents an individual app, and the columns provide various details about each app. Here's a brief description of each column, often referred to as a data dictionary:

#1. App: The name of the app.
#2. Category: The category under which the app falls.
#3. Rating: The average user rating of the app (on a scale of 1 to 5).
#4. Reviews: The number of user reviews.
#5. Size: The size of the app.
#6. Installs: The number of times the app has been installed.
#7. Type: Whether the app is free or paid.
#8. Price: The price of the app (if it is paid).
#9. Content Rating: The age group for which the app is appropriate.
#10. Genres: An additional category description.
#11. Last Updated: The date when the app was last updated on the Play Store.
#12. Current Ver: The current version of the app.
#13. Android Ver: The minimum Android version required to run the app.
```

```{r}
#2. Clean your data. Cleaning involves any task that prepares the dataset for analysis.

# Data Cleaning
# Renaming columns
colnames(google_play_store) <- c('app_name', 'category', 'rating', 'reviews', 'size', 'installs', 'type', 'price', 'content_rating', 'genres', 'last_updated', 'current_version', 'android_version')

# Filling or removing NA values
google_play_store$rating[is.na(google_play_store$rating)] <- mean(google_play_store$rating, na.rm = TRUE)
google_play_store$type[is.na(google_play_store$type)] <- 'Free'
google_play_store$content_rating[is.na(google_play_store$content_rating)] <- 'Everyone'

# Correcting data types
google_play_store$reviews <- as.integer(google_play_store$reviews)
google_play_store$installs <- as.integer(gsub('[+,]', '', google_play_store$installs))
google_play_store$price <- as.numeric(gsub('\\$', '', google_play_store$price))
google_play_store

```

#The data has been successfully cleaned and is now ready for analysis. Here's a summary of the cleaning steps performed:

#1. Renamed columns for easier access.
#2. Managed NAs by filling or dropping them.
#3. Corrected data types for several columns, such as 'reviews', 'installs', and 'price'.
#4. Removed a problematic row with a non-integer 'reviews' value.

```{r}
#3. Determine descriptive statistics for interesting variables.

# Assuming google_play_store is your cleaned dataframe

# Selecting interesting variables
interesting_vars <- google_play_store %>% select(rating, reviews, installs, price)

# Determine descriptive statistics for the interesting variables
summary(interesting_vars)

```
```{r}
#4. Produce visualizations from the raw data that identify and highlight interesting aspects. These can include bar charts, histograms, line graphs, scatter plots, etc. Be sure the chosen graph best represents the information.

# Plot 1: Distribution of App Ratings
plot1 <- ggplot(google_play_store, aes(x=rating)) +
  geom_histogram(binwidth=0.1, fill="blue") +
  ggtitle("Distribution of App Ratings") +
  xlab("Rating") +
  ylab("Frequency")
plot1

```
```{r}

# Plot 2: Distribution of App Reviews (Log Scale)
plot2 <- ggplot(google_play_store %>% filter(is.finite(reviews)), aes(x=reviews)) +
  geom_histogram(binwidth=10000, fill="blue", aes(y = after_stat(count) + 1)) +  # Using after_stat(count) to avoid zero counts
  scale_y_log10() +
  ggtitle("Distribution of App Reviews (Log Scale)") +
  xlab("Reviews") +
  ylab("Frequency (Log Scale)")
plot2

```
```{r}
# Plot 3: Distribution of App Installs (Log Scale),

# Filter the data to only include rows where 'installs' is below a certain threshold
# This threshold can be adjusted based on the data distribution
google_play_store_filtered <- google_play_store %>% filter(installs < 1e6, installs > 0, !is.na(installs))

# Now plot this filtered data
plot3 <- ggplot(google_play_store_filtered, aes(x=installs)) +
  geom_histogram(binwidth=10000, fill="blue") +
  ggtitle("Distribution of App Installs (Filtered)") +
  xlab("Installs") +
  ylab("Frequency")
  
plot3

```
```{r}
# Plot 4: Distribution of App Prices (Log Scale for Paid Apps)

# Filter the data to include only paid apps
google_play_store <- subset(google_play_store, price > 0 & !is.na(price))

# Calculate the frequency of each price bin
freq_data <- as.data.frame(table(cut(filtered_google_play_store$price, breaks=seq(0, max(filtered_google_play_store$price)+1, by=1))))

# Remove bins with zero frequency
freq_data <- freq_data[freq_data$Freq > 0,]

# Plot 4: Distribution of App Prices (Log Scale for Paid Apps)
price_plot <- ggplot(freq_data, aes(x=Var1, y=Freq)) +
  geom_bar(stat="identity", fill="blue") +
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  ggtitle("Distribution of App Prices (Log Scale for Paid Apps)") +
  xlab("Price") +
  ylab("Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Show the plot
price_plot


```
# Part II – Expanding

```{r}
# 1. Create new variables that better capture the data you want to report. For example, if the data shows yearly sales by month, you might calculate the month-to-month increase or decrease in sales as a new column.

# Assuming df is your cleaned dataframe

# Creating a new variable 'reviews_per_install' to capture how many reviews an app gets per install
google_play_store <- google_play_store %>%
  mutate(reviews_per_install = reviews / installs)

# Creating a new variable 'is_paid' to easily filter between free and paid apps
google_play_store <- google_play_store %>%
  mutate(is_paid = ifelse(type == 'Paid', 1, 0))

# Creating a new variable 'price_category' to categorize apps based on their price range
bins <- c(0, 1, 5, 10, 50, 100, 500)
labels <- c('Free', 'Very Cheap', 'Cheap', 'Moderate', 'Expensive', 'Very Expensive')
google_play_store <- google_play_store %>%
  mutate(price_category = cut(price, breaks = bins, labels = labels, right = FALSE))

# Checking the data after creating new variables
head(select(google_play_store, reviews_per_install, is_paid, price_category))
google_play_store

```
#New variables have been created to better capture interesting aspects of the data:

#reviews_per_install: This variable captures how many reviews there are per install for each app. It could be a measure of how much users interact or engage with the app.

#is_paid: A binary variable to easily identify if an app is free (0) or paid (1).

#year_last_updated: This variable captures the year the app was last updated.

#month_last_updated: This variable captures the month the app was last updated.

```{r}
# 2. Group, summarize, rank, arrange, count, or perform any other useful operations to create new data frames that provide access to different views of the data.

# 1. Grouping by 'category' and summarizing to get the mean values for 'rating', 'reviews', 'installs', and 'reviews_per_install'
grouped_by_category <- google_play_store %>%
  group_by(category) %>%
  summarise(mean_rating = mean(rating, na.rm = TRUE),
            mean_reviews = mean(reviews, na.rm = TRUE),
            mean_installs = mean(installs, na.rm = TRUE),
            mean_reviews_per_install = mean(reviews_per_install, na.rm = TRUE),
            .groups = "drop")

# 2. Ranking categories by mean rating
grouped_by_category <- grouped_by_category %>%
  mutate(rank_by_rating = rank(desc(mean_rating)))

# 3. Ranking categories by mean number of installs
grouped_by_category <- grouped_by_category %>%
  mutate(rank_by_installs = rank(desc(mean_installs)))

# 4. Ranking categories by mean number of reviews per install
grouped_by_category <- grouped_by_category %>%
  mutate(rank_by_reviews_per_install = rank(desc(mean_reviews_per_install)))

# 5. Counting the number of free and paid apps in each category
count_by_type <- google_play_store %>%
  group_by(category, is_paid) %>%
  summarise(count = n(), .groups = "drop") %>%
  spread(key = is_paid, value = count, fill = 0) %>%
  rename(free_apps = `0`, paid_apps = `1`)

# Merging the summarized data frames to provide a comprehensive view
merged_summary <- left_join(grouped_by_category, count_by_type, by = "category")

# Displaying the new data frames
arrange(merged_summary, rank_by_rating)
arrange(merged_summary, rank_by_installs)

```
#Two new data frames have been created to provide different views of the data:

#1. By Rating: The table is sorted by the average rating of each app category. The "Education" category leads with the highest average rating, followed by "Events" and "Art and Design".

#2. By Installs: The table is sorted by the average number of installs in each category. The "Communication" category has the most installs on average, followed by "Social" and "Video Players".

#These summaries include rankings and counts for free and paid apps, offering a comprehensive view of different aspects of the app categories.

```{r}
# 3. Extract the most interesting data results and produce visualizations that best communicate these results.

# Assuming merged_summary is your data frame containing all the metrics and rankings

# 1. Visualizing the top 10 categories by mean rating
top_categories_by_rating <- merged_summary %>% 
  arrange(rank_by_rating) %>% 
  head(10)

chart1 <- ggplot(top_categories_by_rating, aes(x=reorder(category, -mean_rating), y=mean_rating)) +
  geom_bar(stat='identity', fill='blue') +
  coord_flip() +
  ggtitle("Top 10 Categories by Mean Rating") +
  xlab("Category") +
  ylab("Mean Rating")

# 2. Visualizing the top 10 categories by mean number of installs
top_categories_by_installs <- merged_summary %>% 
  arrange(rank_by_installs) %>% 
  head(10)

chart2 <- ggplot(top_categories_by_installs, aes(x=reorder(category, -mean_installs), y=mean_installs)) +
  geom_bar(stat='identity', fill='blue') +
  coord_flip() +
  ggtitle("Top 10 Categories by Mean Number of Installs") +
  xlab("Category") +
  ylab("Mean Number of Installs")

# 3. Visualizing the top 10 categories by mean reviews per install
top_categories_by_reviews_per_install <- merged_summary %>% 
  arrange(rank_by_reviews_per_install) %>% 
  head(10)

chart3 <- ggplot(top_categories_by_reviews_per_install, aes(x=reorder(category, -mean_reviews_per_install), y=mean_reviews_per_install)) +
  geom_bar(stat='identity', fill='blue') +
  coord_flip() +
  ggtitle("Top 10 Categories by Mean Reviews per Install") +
  xlab("Category") +
  ylab("Mean Reviews per Install")

# Displaying the plots
chart1

```
```{r}
chart2
```

```{r}
chart3
```
#The visualizations highlight the most interesting insights from the Google Play Store Apps data:

#1. Top 10 Categories by Mean Rating: The "Education" category leads with the highest average rating, followed by "Events" and "Art and Design".

#2. Top 10 Categories by Mean Number of Installs: The "Communication" category has the most installs on average, followed by "Social" and "Video Players".

#3. Top 10 Categories by Mean Reviews per Install: The "Game" category has the highest number of reviews per install, indicating high user engagement. This is followed by "Social" and "Dating" categories.

#These visualizations provide a comprehensive view of the performance and engagement metrics across different app categories.


# Part III – Communicating

```{r}
#1. Report what you have learned about your data. Identify 3–5 observations or follow- up questions that you could explore in the future.

#Observations:
#1. Highly Rated Categories: The "Education" category has the highest average rating, suggesting that educational apps generally meet user expectations.

#2. Most Installed Categories: Apps in the "Communication" category have the highest average number of installs, indicating a strong user demand for these types of apps.

#3. User Engagement: The "Game" category has the highest average reviews per install, signifying high user engagement and interaction.

#4. Free vs Paid Apps: The majority of apps in the dataset are free, but this does not necessarily correlate with higher ratings.

#5. Review Patterns: Categories like "Social" and "Dating" also show high user engagement, as indicated by the reviews per install metric.

#Follow-Up Questions for Future Exploration:
#1. User Retention: How does the number of reviews correlate with user retention metrics, if such data were available?

#2. Seasonal Trends: Are there seasonal trends in app installs or ratings? For example, do educational apps see more installs during the school season?

#3. Monetization: How does the type of monetization (ads, in-app purchases, etc.) affect user ratings and engagement?

#4. Content Rating: How does the content rating (e.g., "Everyone", "Teen") affect user engagement and ratings?

#5. App Updates: Is there a correlation between the frequency of app updates and user ratings or reviews?

```



